#Inference Project

##Question 1

Here's the simulation code.

```{r}
lambda=.2
m=1000
n=40
means=numeric(m)
set.seed(1)
for (i in 1:m){
means[i]=mean(rexp(n,lambda))
}
```

###Compare Theoretical and Simulation Means

The simulation median is shown below:
```{r}
median(means)
```

The theoretical median is 1/lambda or 5, so we can see that the simulation and theoretical values match up well.

###Compare variance:

The simulation variance is calculated:
```{r}
var(means)
```

The theoretical population variance is (1/lambda)^2=25.  The theoretical sample variance is popluation variance/sqrt(n) or 25/40=.625.  As we can, see the variances match up well.

###Show distribution is relatively normal:

I plot the histogram of the simulation and create a normal distribution that should theoretically match with the histogram.  We see that the simulation matches theory well.

```{r}
x=seq(2,8,.1)
hist(means,20)
lines(x,1000/5*dnorm(x,mean=mean(means),sd=sd(means)),col="red")
```

### coverage of confidence interval

We look at the confidence interval:
```{r}
range=c(-1,1)*1.96*(1/lambda)/sqrt(n)+mean(means)
coverage=sum(means>=range[1] & means<=range[2])
coverage/m*100
```

Notice that the coverage matches the theoretical coverage of 95% well.

##Problem 2:

Let's load the data first.

```{r}
data(ToothGrowth)
data=ToothGrowth
```

###Exploratory Analysis:

Below, is a histogram of lengths for various dosages and supplements.

```{r}
library(ggplot2)
ggplot(data,aes(x=len))+geom_histogram(binwidth=2)+facet_grid(dose~supp)+ggtitle("Length histogram for different dosages and supp")
```

Below, we see the means of the data for various dosages and supplements.
```{r}
tapply(data$len,list(data$supp,data$dose),mean)
```

From the exploratory analysis, we see that length tends to increase with dosage, and OJ seems to have a larger effect than VC.

###Summary of data
```{r}
summary(data)
```

We see taht there are 2 types of supplements, with 30 data points each.  W also see taht the mean length is around 19.  Also, dose ranges from .5 to 2.

###OJ versus VC Inference Analysis:

We will look at OJ versus VC for each dose: .5,1, and 2. We will calculate the two-tail p-value and 95% 2-tail confidence interval for each dose.

```{r}
OJ=subset(data$len,data$supp=="OJ"&data$dose==.5)
VC=subset(data$len,data$supp=="VC"&data$dose==.5)
p_val.5=t.test(OJ,VC,var.equal=FALSE)$p.value
conf.int.5=t.test(OJ,VC,var.equal=FALSE)$conf.int
p_val.5
conf.int.5

OJ=subset(data$len,data$supp=="OJ"&data$dose==1)
VC=subset(data$len,data$supp=="VC"&data$dose==1)
p_val1=t.test(OJ,VC,var.equal=FALSE)$p.value
conf.int1=t.test(OJ,VC,var.equal=FALSE)$conf.int
p_val1
conf.int1

OJ=subset(data$len,data$supp=="OJ"&data$dose==2)
VC=subset(data$len,data$supp=="VC"&data$dose==2)
p_val2=t.test(OJ,VC,var.equal=FALSE)$p.value
conf.int2=t.test(OJ,VC,var.equal=FALSE)$conf.int
p_val2
conf.int2
```

##Results and Analysis:

Given alpha=.05 for a two-tailed p value test, we see that OJ increases length more than VC for both .5 and 1 dosages.  For dose 2, we accept the null hypothesis.  We see that the 95% confidence intervals also confirm this in that the intervals for .5 and 1 doses are completely positive whereas dose 2 ranges from negative to positive values. We can also account for the multiple hypothesis tests by dividing alpha by 3, but this would still have the first 2 p-values within our alpha limit, so it doesn't change our results.
